Data-Driven Modelling \& \\ Forecasting of Chaotic Systems

### 2022-01-16 draft_16Jan/draft5.tex
Many graphs added and comments added. Edits of Ch.5. Starting to prep to add Grammarly back in

\begin{abstract}
  Collecting measurements from underlying artificial, natural, and physical systems through observables to realise a dynamical model with greater descriptive power is a well-established practice. Physical systems can often be modelled by instance with differential equations derived from the laws of physics, if we knew the laws.  However, we often do not fully understand the many complexities of the more intricate systems we interact with on a daily basis.
  In this project we start with the hypothesis that an discrete-time dynamical systems, particularly those which are chaotic generates the data. The general learning problem is then formulated and a well-known approach to the problem, alongside its practical limitations are considered.
  We then introduce the notion of driven dynamical systems and establish the theoretical existence of dynamical systems possessing certain properties that guarantee them to be topologically conjugate ('dynamically equivalent') to the dynamics of the underlying system from which measurements were taken. The methodology is practically implemented by making use of a recurrent neural network and using deep learning methods to train on the recurrent neural network states to show that we may obtain long-term consistent predictions for simple physical systems such as the double pendulum as well as a few other attractors exhibiting chaotic behaviour. The machine learning algorithms for modelling the double pendulum thus far have not been successful when it exhibits chaos.

  \end{abstract}

@@

  In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$.
  This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.  So $T$ has sensitive dependence on initial conditions.

@@

\ednote{M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity. So please edit accordingly. B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough? }
When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent''. Specifically, they are in one-to-one correspondence with one another. However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$ \cite{de2013elements}. When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}. In essence, an extension (e.g., \cite{de2013elements}) is a larger system capturing all of the important dynamics of its factor.

@@

First we establish the method of Takens delay embedding. 
In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{Should this manifold be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which will be formalised below.
We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e., we observe  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$. The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$. 
Consider for example a thermometer erected to measure the ambient temperature in a local village. This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the present weather of the surrounding area. Even more than that though- it is measuring a miniscule part of the global weather system.

@@

However, the true state $u$ of a system is seldom, if ever, fully known. In almost all cases we can at most insert probes into a system to obtain partial information by means of the measurements taken. Moreover, the process of taking a measurement itself introduces 2 additional aspects which complicate the problem: \ednote{M: Maybe replace difficulty by saying that we need to consider two aspects. B: Perhaps this edit makes more sense?}
\vspace{-8mm}
\begin{enumerate}[noitemsep, label=\roman*.]
  \item A series of measurements over a specific time-interval is inherently a discretisation procedure of the underlying continuous-time dynamical system. (Hence why we restricted our attention to discrete-time systems in the preceding section)
  \item A probe will never be fully accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
\end{enumerate}

@@

\section{Practical Limitations to Takens' Theorem}\label{sect_takenslimits}

@@

The Takens embedding theorem does not guarantee global dissipativity.
\ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.  Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}

\ednote{Include note on learning failure because of sensitivity as in graph commented out?}
% \begin{figure}[ht]
%   \includegraphics[scale=0.3]{_learningfailure.eps}
%   \centering
%   \caption{Illustration of a system having failed to learn the map}
%   \label{fig:supp_learning_failure}
% \end{figure}

@@

\begin{Definition}
  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP})]\label{Dfn_NDS}\rm
  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous. Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space. 
\end{Definition}

@@

\begin{Example} \rm \label{ex_halfux}
  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
\end{Example}

@@

As the solutions to a driven system are often considered to be an important entity \cite{M: edited the sentence here}, we next identify a subspace $X_U$ of $X$ that contains all possible solutions. To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.

@@

The reachable set itself can defined independently from the fact that $g$ does or does not have the USP. We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$. \textbf{(Cite)}.
The set $X_n(\overline{u})$ is exactly the set of all points $x$ that have some entire solution $\Psi$ passes through it at time $n$. \ednote{B: Changed sentence for better readability - still understandable?}

We can give an alternate formulation of the set $X_n(\overline{u})$:
suppose, for some fixed input 
$\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states
\begin{equation} \label{eqn_association}
X_{n,i}(\overline{u}) := g_{n-1} \circ \cdots g_{i+1} \circ g_i(X).
\end{equation}

As the image of a finite composition of continuous maps, the set $X_{n,i}(\overline{u})$ is compact whenever $X$ is compact and is also nonempty.
Further, $X_{n,i}(\overline{u}) \supset X_{n,i-1}(\overline{u})$. Hence $\bigcap_{i<n}
X_{n,i}(\overline{u})$ is a nested intersection of closed nonempty subsets, and
whenever $X$ is compact, the intersection is nonempty.  \ednote{B: Should I cite here?}

\begin{Definition}
  [\bf Topological Contraction]\label{Dfn_TopContr}\rm
  We say $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$. 
\end{Definition}

It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n(\overline{u})$ that we had defined earlier \ednote{(Insert citations.)}. This will lead to a result (for. e.g., \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}

@@

Thus far it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies certain properties such as uniqueness. 
Should the driven system be compact, existence follows immediately.

\begin{Theorem}\label{Thm_CompactExistence}
 Let $g$ be a driven system.  If $X$ is compact, then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
\end{Theorem}
{\bf Proof.}  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}

@@

\ednote{B: RC introduced a bit abruptly. Adjust.} Reservoir Computing(RC) refers to a machine learning approach in which input is mapped onto a different, but higher dimensional, space through a system called the reservoir after which a readout measuring the reservoir state is trained.
In the majority of Reservoir Computing(RC) literature \textbf{(Cite)}, a notion of forgetting the states of the reservoir asymptotically is often used.  
Concepts like the echo state property(ESP) and  fading memory \cite{boyd1985fading} are two of these.  
If $g$ possesses the ESP (equivalent in our context to the USP \textbf{(Cite)}), we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e. there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo,Manju_2020}.
% \cite{Manju_ESP}.
Beyond forgetting the past states, the ESP is also often discussed in terms of a stability property that plays a key role in the robustness of recurrent neural networks(RNN).
The concept, termed input-related stability~\cite{manjunath2020stability}, refers the phenomenon of nearby inputs creating responses (solutions) close to one another.

\section{Establishing Conjugacies}

@@

But does such a function $h$ for the above driven system $g$ above exist? Whenever $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
Regrettably, such a continuous mapping $h$ is not guaranteed to exist \cite{M: Have edited here} when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.

@@

The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$. 
The continuity of $H^{-1}$ follows from the fact that $H$ is itself a continuous and surjective mapping of a compact space $\overleftarrow{U}$ in a Hausdorff space.
and we refer to the map $H$ as a \emph{causal embedding}.

@@

The inverse-limit space is well-defined since $T : U \to{U}$ is surjective by assumption.We now get to the concept of embedding the inverse limit-space of the dynamical system in $X \times X$.

\begin{Definition} \rm
	We say a driven system $g$ \emph{causally embeds} a dynamical system $(U,T)$ if it satisfies the two properties: (i) a universal semi-conjugacy exists and  (ii)  $H_2(\cev{u}) := (h(r\cev{u}),h(\cev{u}))$ embeds the inverse-limit space $\widehat{U}_T$ in $X \times X$. \end{Definition}

It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$.

@@

The following theorem establishes the existence of a well-defined map $G_T$ describing the SDD of the system above.

@@

\begin{Definition}
  [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
  A driven system $g$ has the uniform attraction property (UAP) if we initialize the driven system
with an arbitrary initial value $y_m \in X$, and the sequence $y_{m+1}, y_{m+2}, y_{m+3},...$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ then approximates an actual solution $\{x_n\}$ uniformly in the sense that given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that $d(x_{n+i}, y_{n+i})<\epsilon$ for all $i\ge 0$, where $\{x_m\}$ is a solution.
\end{Definition}

@@

%\begin{Theorem}\label{thm_usp-contr-uap}  The following statements are equivalent:  \vspace{-8mm}  \begin{enumerate}[noitemsep, label=\roman*.]    \item $g$ has the USP.    \item $g$ is a topological contraction.    \item $g$ has the UAP.  \end{enumerate\end{Theorem} {\bf Proof.}  May be found in \cite[Th.1]{Manju_Nonlinearity}

@@

In this section we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$. We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient. 
Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum. We also consider other chaotic attractors.

\section{Implementing $G_T$}
The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}. To reach this point, one must first learn $\Gamma$.
After first normalising the data to ensure it lies within the interval $|x| < 1$ and has zero mean value, we continue by driving the system with external input through the function $g$.

From \eqref{eqn_driving} we know that $ g(u,x)=(1-a)x+a{\cdot}\tanh(Au+\alpha{a}Bx)$. Recall that $g$ is SI-invertible. Adopting the terminology of \cite{manjunath2021universal}, $A$ is called the “input matrix” and $B$ the “reservoir matrix”.

We make use of a Recurrent Neural Network (RNN) in our implementation. RNN's are widely used in the literature for discrete-time processing and are considered to be discrete-time version of continuous dynamical systems modelling a network of neurons. (\textbf{Citation.})
A complete discussion of RNN's is outside the scope of this thesis, so we present only the details necessary to the current discussion. For information, (\textbf{Cite, Cite, Cite}) may be consulted. We implement our RNN in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library.

If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices. Inputs $u_n$ of dimension $K < N$ are padded by zeroes to obtain an input of dimension $N$. As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution’s elements in a uniform manner thanks to the UAP. \textbf{(Citation)}

Thirdly we prepare the network. Utilising a network $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$). Training is done using the Adam Optimiser where the MSE loss function is optimised. The system is trained several times(in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run). 
Additional parameters such as the number of network-layers, neurons per layer and training-length are presented case-by-case below.

@@

\section{Changes in this project to previous work of~\cite{manjunath2021universal}}

To understand the first and most drastic\ednote{change word} change, we recall the the conjugacy in~\ref{fig:inverse_limsystem}:

\begin{equation*}
  %  \[ 
      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
      \everypsbox{\scriptstyle}
      \begin{psmatrix}
      \widehat{U}_T  && \widehat{U}_T \\
      Y_T && Y_T \\
      %%%
     %  \ncline{1,1}{1,2}^{\widehat{T}} \ncline{1,1}{2,1} <{h}
     %  \ncline{1,2}{2,2} > {H_2}
     %  \ncline{2,1}{2,2}^{G_T}
      \end{psmatrix}
    %  \]
  \end{equation*}

  We recall that $G_T$ is topologically conjugate to $\widehat{T}$, a result which we have already made use of.

  The authors of~\cite{manjunath2021universal} did not, however, consider the relationship between the inverse-limit space $\widehat{U}_T$ and the input $u_n=\theta(w_n)$ (where $\theta$ is our observable and $w_n$ the complete input), a concept we now explore.

>>>>

\begin{abstract}
  Collecting measurements from underlying artificial, natural, and physical systems through observables to realise a dynamical model with greater descriptive power is a well-established practice. Physical systems can often be modelled by instance with differential equations derived from the laws of physics, if we knew the laws.  However, we often do not fully understand the many complexities of the more intricate systems we interact with on a daily basis.
  In this project we start with the hypothesis that an discrete-time dynamical systems, particularly those which are chaotic generates the data. The general learning problem is then formulated and a well-known approach to the problem, alongside its practical limitations are considered.
  We then introduce the notion of driven dynamical systems and establish the theoretical existence of dynamical systems possessing certain properties that guarantee them to be topologically conjugate ('dynamically equivalent') to the dynamics of the underlying system from which measurements were taken. The methodology is practically implemented by making use of a recurrent neural network and using deep learning methods to train on the recurrent neural network states to show that we may obtain long-term consistent predictions for simple physical systems such as the double pendulum as well as a few other attractors exhibiting chaotic behaviour. The machine learning algorithms for modelling the double pendulum thus far have not been successful when it exhibits chaos. 
  \end{abstract}

@@

  In general, the graph of  $T^k$ contains $2^{k-1}$ tents.
  Given any point $u$ and an open neighborhood $W \subset[0,1]$ of $u$, we can find $k\in\mathbb{Z}$ large enough so that $T^k(W) = [0,1]$ because we can accommodate a tent whose base is contained in $W$. % \ednote{B: This last phrase doesn't make sense to me. M: The number of tents grow with $k$ and hence this is obvious; I explained in one of our meetings. Does it make sense?} 
  This implies the existence of two points $w_1$ and $w_2$ in $W$ so that $d(T^kw_1,T^kw_2)= 1$, and by the triangle inequality  at least one of the inequalities $d(T^ku,T^kw_1)> \delta$ or $d(T^ku,T^kw_2)> \delta$ hold when $\delta\in (0,\frac{1}{2})$.  So $T$ has sensitive dependence on initial conditions.

@@

\ednote{B: Does defining explicitly the domain and codomain of $\phi$ make it clear enough?  M: Saying $\phi$ is then a semi-conjugacy between $T$ and $S$ could be ambiguous as it does not specify which system is on the top in the commutativity. So please edit accordingly.}
When $S$ is conjugate to $T$, the dynamics of the two systems are in some way ''dynamically equivalent''. Specifically, they are in one-to-one correspondence with one another. However when $S$ is semi-conjugate to $T$ with $\phi:U\to{V}$ a many-to-one mapping, the dynamics on $V$ provide merely a coarse-grained description of the dynamics on $U$ \cite{de2013elements}. When $S$ is semi-conjugate to $T$, it is also common to call $S$ a \emph{factor of $T$}, or conversely that $T$ is an \emph{extension of $S$}. In essence, an extension (e.g., \cite{de2013elements}) is a larger system capturing all of the important dynamics of its factor.

@@

First we establish the method of Takens delay embedding. 
In this method, one considers a discrete-time dynamical system defined on a 'nice' space (a smooth manifold)\ednote{B: Should this be further explained?} that can be obtained as time-$K$ map of a continuous time-dynamical system, concepts which will be formalised below.
We make observations from such a system, i.e., we observe the evolution of $\theta(w_0)$ (where $w_0$ represents the initial state) i.e., we observe  a finite set of values of $u_n :=  \theta(w_n) = \theta(Tw_{n-1})$. The sequence $\{u_n\}$ represents a scalar time-series and intuitively $\theta$ may be thought to represent a probe inserted into a larger system which is itself only measuring/extracting a small part of the greater system state at time $t=n$. 
Consider for example a thermometer erected to measure the ambient temperature in a local village. This measurement function, the thermometer, is capturing only a single aspect, the temperature, of a much grander dynamical system entailing the present weather of the surrounding area. Even more than that though- it is measuring a miniscule part of the global weather system.

@@

However, the true state $u$ of a system is seldom, if ever, fully known. In almost all cases we can at most insert probes into a system to obtain partial information by means of the measurements taken. Moreover, the process of taking a measurement itself introduces 2 additional aspects which complicate the problem: \ednote{B: Perhaps this edit makes more sense? M: Maybe replace difficulty by saying that we need to consider two aspects. }
\vspace{-8mm}
\begin{enumerate}[noitemsep, label=\roman*.]
  \item A series of measurements over a specific time-interval is inherently a discretisation procedure of the underlying continuous-time dynamical system. (Hence why we restricted our attention to discrete-time systems in the preceding section)
  \item A probe will never be fully accurate, and so the act of measurement introduces a certain measure of numerical noise/inaccuracy.
\end{enumerate}

@@

\section{Practical Limitations to Takens' Theorem}

@@

% The Takens embedding theorem does not guarantee global dissipativity.
% \ednote{B:Links with~\ref{subs_LearnGamma} in discussion of dissipativity of $\Gamma$.  Sir, I'm not sure how to write this here - I think it a beneficial point to make so as to set the stage for the later discussion in ~\ref{subs_LearnGamma}, but I am unsure there as well.}

@@

\begin{Definition}
  [\bf Nonautonomous Dynamical System (adopted from~\cite{Manju_ESP}))] \label{Dfn_NDS}\rm
  A nonautonomous dynamical system (NDS) on a space $X$ is simply a dynamical system comprising a family of maps $\{g_n\}_{n \in \mathbb{Z}}$ where $g_n(\ldotp):=g(u_n,\ldotp):X\to{X}$ is continuous. Each $g_n$ arises as a consequence of an input $u_n$ from the input space $U$, a topological space. 
\end{Definition}

@@

\begin{Example} \rm \label{ex_halfux}
  The only solution  $\{x_n\}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence , say $u_n=0.5$. 
  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
\end{Example}
% \ednote{Nicely written! Thank you}

@@

As the solutions to a driven system are often considered to be an important entity, we next identify a subspace $X_U$ of $X$ that contains all possible solutions. To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.

@@

The reachable set itself can defined independently from the fact that $g$ does or does not have the ESP/USP. We note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$. \textbf{(Cite)}.

The set $X_n(\overline{u})$ is exactly the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$. 
Now suppose, we for some fixed input 
$\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states

\begin{equation} \label{eqn_association}
X_{n,i}(\overline{u}) := g_{n-1} \circ \cdots g_{i+1} \circ g_i(X).
\end{equation}

The set $X_{n,i}(\overline{u})$ being the image of a finite composition of
continuous maps is compact whenever $X$ is compact. Also $X_{n,i}(\overline{u})$ is
nonempty.  Further, $X_{n,i}(\overline{u}) \supset X_{n,i-1}$. Hence $\bigcap_{i<n}
X_{n,i}(\overline{u})$ is a nested intersection of closed nonempty subsets, and
whenever $X$ is compact, the intersection is nonempty. We say $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$.

It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier. This will lead to a result (for. e.g., \cite{manjunath2013echo}) that  \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}

% \begin{Definition}
%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
% \end{Definition}

@@

Thus far it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies certain properties such as uniqueness. 
Should the driven system be compact, existence follows immediately as shown in the following result.

\begin{Theorem}\label{Thm_CompactExistence}
 Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
\end{Theorem}
\begin{proof}
  May be found in \cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
\end{proof}

@@

In the majority of Reservoir Computing(RC) literature (a machine learning approach) where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.

% \ednote{How should I explain/qualify RC better here? It's mentioned without really explaining. M: Have put it in one sentence; just to let you know people use optical devices to map signal into a ``reservoir". B: Thanks!}, 
Concepts like the ESP,  fading memory \cite{boyd1985fading} are some of these notions.  
If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e. there is only ever the possibility of a single reachable state at a given instance of time \cite{jaeger2001echo,Manju_2020}.
% Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described. A more detailed presentation may be found in~\cite{jaeger2001echo}.

% We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks(RNN), a component of the implementation discussed in~\ref{ch5}.

\section{Conjugacies}

@@

But does such a function $h$ for the above driven system $g$ above exist? Whenever $g$ has the USP and $\Psi(u)=\{x_n\}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above \ref{Scomm_h}.
Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP \cite[Lemma 5]{Manju_Nonlinearity}.
Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.

@@

The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$. 
The continuity of $H^{-1}$ follows from the fact that $H$ is itself a continuous and surjective mapping of a compact space $\overleftarrow{U}$ in a Hausdorff space.
and we refer to the map $H$ as a \emph{causal embedding}. 
\ednote{B: edited the sentence here}

@@

The inverse-limit space is well-defined since $T : U \to{U}$ is surjective by assumption.
We now get to the concept of embedding the inverse limit-space of the dynamical system in $X \times X$. 
\begin{Definition} \rm
	We say a driven system $g$ \emph{causally embeds} a dynamical system $(U,T)$ if it satisfies the two properties: (i) a universal semi-conjugacy exists and  (ii)  $H_2(\cev{u}) := (h(r\cev{u}),h(\cev{u}))$ embeds the inverse-limit space $\widehat{U}_T$ in $X \times X$. \end{Definition}

  It is imperative to take note of a subtlety here: we refer to the \emph{map} $H$ as a (causal) embedding, whereas the driven \emph{system} $g$ (causally) embeds another \emph{system} $(U,T)$. \ednote{B: inserted line}

@@

The following theorem establishes the existence of a well-defined map $G_T$ describing the single-delay dynamics(SDD) of the system above.

@@

\begin{Definition}
  [\bf Uniform Attraction Property]\label{Dfn_UAP}\rm
  A driven system $g$ has the Uniform Attraction Property (UAP) if we initialize the driven system
with an arbitrary initial value $y_m \in X$, and the sequence $y_{m+1}, y_{m+2}, y_{m+3},...$ satisfying $y_{k+1}= g(u_k,y_k)$ for $k \geq m$ then approximates an actual solution $\{x_n\}$ uniformly in the sense that given $\epsilon>0$ (independent of $y_m$) there is an integer $n$ so that $d(x_{n+i}, y_{n+i})<\epsilon$ for all $i\ge 0$, where $\{x_m\}$ is a solution.
\end{Definition}

@@

%\begin{Theorem}\label{thm_usp-contr-uap}  The following statements are equivalent:  \vspace{-8mm}  \begin{enumerate}[noitemsep, label=\roman*.]    \item $g$ has the USP.    \item $g$ is a topological contraction.    \item $g$ has the UAP.  \end{enumerate\end{Theorem}{\bf Proof.}  May be found in \cite[Th.1]{Manju_Nonlinearity}

@@

In this chapter we turn to the implementation of the function $G_T$ by discussing a method to learn the function $\Gamma$. We apply our methodology to some chaotic attractors. Most notably, we consider the reconstruction of the chaotic attractor of the double pendulum - an  attractor that has not been successfully reconstructed because of its intricate structure. In fact the previous data-driven approaches also have not been successful in building a model from the double pendulum.

%We then indicate where this implementation differs from previous work done by (\cite{manjunath2021universal}) and empirically show that the SDD of $G_T$ are less functionally complex than $\Phi_{k, \theta}$ or $T$ by considering the Pearson correlation coefficient. Finally, we present and discuss some numerical results for a focus system of this project: the double pendulum. We also consider other chaotic attractors.

\section{Implementing $G_T$}
The map $G_T$ is implemented by applying the functions $\Gamma$, $g$ and the projection function $\pi_2$ in the relevant order as described by the equations~\ref{eqns_from_data}. To reach this point, one must first learn $\Gamma$.
After first normalising the data to ensure that it has zero mean value and lies within the range $[-1,1]$ and required zero-padding \ednote{M: this zero padding is something that we need to tell the examiner}, we continue by driving the system with external input through the function $g$.

From \eqref{eqn_driving} we know that $g(u,x) = (1-a) x + a \tanh(Au + \alpha B x)$. Recall that $g$ is SI-invertible. We refer to $A$ as the input matrix and $B$ the reservoir matrix.

We make use of a Recurrent Neural Network (RNN) in our implementation to project the data on to a higher dimensional space, but we do not train the RNN in the conventional manner. \ednote{B: Should we say what the conventional manner is?} A few words about the RNNs: RNNs are widely used in the literature for temporal processing of data as a type of deep learning technique. They take the form of equations considered to be discrete-time version of continuous dynamical systems modelling a neural network of neurons \cite{jaeger2001echo}.

We implement a feedforward network, (a network with no memory, unlike the RNN)\ednote{B: Added so that FFN is in more context} required to learn $\Gamma$ in the \emph{Python} programming language and making use of the \emph{Scipy} library, alongside the \emph{Keras} package in the \emph{Tensorflow} library \ednote{B: Is this something you want me to write in? M: Note! We do not implement the RNN, but the FN that learns Gamma through the Tensorflow. We called the RNN + FN set up a RCN in Adriaan's paper. FYI: FN's work only if there is a definite relationship between the $u_n$ and $u_{n+1}$ which is rare in temporal data.
RNNs with ESP work when there is a definite relationship between $(\ldots,u_{n-1},u_n)$ and $u_{n+1}$ and is suited for temporal processing.}.

If the RNN we will be using in the next step has $N$ neurons, then $A$ and $B$ are each $N\times{N}$ matrices. Inputs $u_n$ of dimension $K < N$ are padded by zeroes (as remarked earlier) to embed the input in a space of dimension $N$. As we do not need the entire left-infinite history of input, an arbitrary initialisation of the system with initial values followed by the driving of the system with the function $g$ means that the generated values will approach the actual solution elements in a uniform manner thanks to the UAP that we had discussed in the preceding chapter.

Thirdly we prepare the network. Utilising a network $N$ neurons (the same value $N$ as in the dimensions of the matrices $A,B$). Learning of $\Gamma$ is done using the Adam Optimiser where the means square error (MSE) loss function is optimised. The system is trained several times (in our case 4) with incrementally smaller learning rates on each run (0.001 and thereafter divided by 10 at each run). 
Additional parameters such as the number of network-layers, neurons per layer and training-length are presented case-by-case below.

@@

\section{Delay Coordinates}
